---
title: "Chapter 2 Exercises"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```

## Getting started with ggplot2

```{r, echo=FALSE}

require(data.table, quietly = TRUE, warn.conflicts = FALSE)
assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

require(dplyr, quietly = TRUE, warn.conflicts = FALSE)
require(ggplot2, quietly = TRUE, warn.conflicts = FALSE)

mpg

```

### Exercises 2.2

> __Question 1__: 

> List five functions that you could use to get more information
    about the `mpg` data set.
    
> __Answer__: Some useful functions are:

> + `summary`: print some summary statistics for each variable  
> + `View`: to see the whole data set in a nice spread-sheet like fashion
> + `str` : get info on the structure of the `mpg` object
> + `dplyr::glimps`: similar to `str` but much tidier
> + `class`: to get its class

```{r ch_2.2.1_1}
glimpse(mpg)
```

> __Question 2__: 

> How can you find out what other datasets are included with ggplot2?

> __Answer__:

```{r ch_2.2.1_3}

pkg.data <- data(package = "ggplot2")
pkgs <- as.data.table(pkg.data$results)[, 3:4]

pkgs

```


> __Question 3__: 

> Apart from the US, most countries use fuel consumption, 
(`fuel consumed over fixed distance`) rather than fuel economy,
(`distance travelled with fixed amount of fuel`).

> How could you convert cty and hwy into the European standard of 1/100km?
    
> __Answer__: 
 

```{r ch_2.2.1_2}

mpgTol100km <- function(milespergallon) {
	GalloLiter <- 3.785411784
	MileKilometer <- 1.609344

	l100km <- (100 * GalloLiter) / (milespergallon * MileKilometer)
	l100km
}


```

> __Question 4__: 

> Which manufacturer has the most models in this dataset?

```{r ch_2_q_4_1}

q4_1 <- mpg %>%
	group_by(manufacturer) %>%
	tally( sort = TRUE)

```


> Which model has the most variations?

```{r ch_2_q_4_2}

q4_2 <- mpg %>%
    group_by(manufacturer) %>%
    transmute("n" = length(unique(model))) %>%
    unique() %>%
    ungroup() %>%
    arrange(desc(n))

```


```{r}

df <- data.frame(rbind(q4_1, q4_2), "Type" = as.factor(rep(c("overall", "unique"), each = 15)))

ggplot(df, aes(x = reorder(as.factor(manufacturer), n), y = n, fill = Type)) +
    geom_bar(width = 0.5, stat = "identity", position = "dodge") +
    theme(panel.grid.major.y = element_blank(),
		panel.grid.minor.y = element_blank()) +
		scale_y_continuous(breaks = seq(0, 40, by = 5)) +
		ggtitle("Number of models per manufacturer") +
		labs(y = "Count", x = "Model") +
		coord_flip()

```

> Does your answer change if you remove the redundant specifications of drive train 
(e.g., "pathfinder 4wd", 'a4 quattro") from the model name?


```{r ch_2_q_4_3}

require(stringr, quietly = TRUE)

unique_model <- mpg %>%
	select(manufacturer, model)

unique_model$model <- str_trim(str_replace_all(unique_model$model, c("quattro" = "", "4wd" = "",
									 "2wd" = "", "awd" = "")))

q4_3 <- unique_model %>%
	group_by(manufacturer) %>%
	transmute("n" = length(unique(model))) %>%
	unique() %>%
	ungroup() %>%
	arrange(desc(n))

df <- data.frame(rbind(q4_1, q4_3), "Type" = as.factor(rep(c("overall", "unique"), each = 15)))

ggplot(df, aes(x = reorder(as.factor(manufacturer), n), y = n, fill = Type)) +
	geom_bar(width = 0.5, stat = "identity", position = "dodge") +
	theme(panel.grid.major.y = element_blank(),
		panel.grid.minor.y = element_blank()) +
		scale_y_continuous(breaks = seq(0, 40, by = 5)) +
		ggtitle("Number of Unique models per manufacturer") +
		labs(y = "Count", x = "Model") +
		coord_flip()

```


```{r}
# rm(list = ls())
```

### Exercises 2.3

> __Question 1__: 

> How would you describe the relationship between cty and hwy? Do you have any
    concerns about drawing conclusions from that plot?
    
> __Answer__:

```{r ch_2.3.1}

ggplot(mpg, aes(hwy, cty)) +
    geom_point()

```

Strongly correlated. I would expect as much from intutition as they are both measurements of fuel economy.

> __Question 2__: 

> What does ggplot( mpg, aes( model, manufacturer)) + geom_point() show? Is it useful? 
How could you modifiy the data to make it more informative?

> __Answer__:


```{r ch_2_3.2_1}

glimpse(mpg)

ggplot(mpg, aes(model, manufacturer)) +
    geom_point()

mpg %>% group_by(model, manufacturer)

```

It shows the manufacturer by model. It is not very useful. A better alternative might be:

```{r ch_2_3.2_2}

df <- mpg %>%
	transmute("man_mod" = paste(manufacturer, model, sep = " "))

ggplot(df, aes(man_mod)) +
	geom_bar()

```

> __Question 3__:

Describe the data, aesthetic mappings and layers used for each of the following plots. 
You'll need to guess a little because you haven't seen all the data sets and functions yet, but use your common sense! 
See if you can predict what the plot will look like before running the code.


```{r ch_2_3.3}

ggplot(mpg, aes(cty, hwy)) + geom_point()
ggplot(diamonds, aes(carat, price)) + geom_point()
ggplot(economics, aes(date, unemploy)) + geom_line()
ggplot(mpg, aes(cty)) + geom_histogram()

```

### Exercises 2.4

> __Question 1__: 

> Experiment with the color, shape and size aesthetics. 
What happens when you map them to continuous values? What about categorical values? 
What happens when you use more than one aesthetic in a plot?
    
> __Answer__:

```{r ch_2.4.1}

# Categorial
ggplot(mpg, aes(cty, displ, colour = class)) +
    geom_point()

# Continuous 
ggplot(mpg, aes(cty, hwy, size = displ)) +
    geom_jitter()

```

> __Question 2__: 

> What happens if you map a continuous variable to shape? Why? What happens if you map trans to shape? Why?
    
> __Answer__:

```{r ch_2.4.2}

ggplot(mpg, aes(cty, hwy, shape = trans)) +
    geom_point()

```

> __Question 2__: 

> How is drive train related to fuel economy? How is drive train related to engine size and class?
    
> __Answer__:

```{r ch_2.4.3}

ggplot(mpg, aes(drv, cty)) +
    geom_boxplot() +
    scale_x_discrete(limits = c("f", "r", "4"),
				   labels = c("Front wheel", "Rear wheel", "Four wheel"))

ggplot(mpg, aes(drv, cty)) +
    geom_violin() +
    scale_x_discrete(limits = c("f", "r", "4"),
				   labels = c("Front wheel", "Rear wheel", "Four wheel"))

ggplot(mpg, aes(reorder(class, displ, FUN = median), displ, colour = drv)) +
	geom_jitter(width = 0.5, height = 0)

```

### Exercises 2.5

> __Question 1__: 

> What happens when you try to facet by a continuous variable like hwy? What about cyl? What's the key difference?

> __Answer__:

The key difference is that faceting by a continious variable leads to the generation of too many plots to be meaningful in analysis. 
The factor based approach works much better.

```{r ch_2_5.1}

ggplot(mpg, aes(displ, cty)) +
	geom_point() +
	facet_wrap(~hwy)

ggplot(mpg, aes(displ, cty)) +
	geom_point() +
	facet_wrap(~cyl)

```

> __Question 2__: 

> Use facetting to explore the 3-way relationship between fuel economy, engine size, and number of cylinders. How does facetting by number of cylinders change your assessement of the relationship between engine size and fuel economy?

> __Answer__:


```{r ch_2_5.2}

ggplot(mpg, aes(displ, cty)) +
	geom_point() +
	geom_smooth(method = "lm") +
	facet_wrap(~cyl)

```

> __Question 3__: 

>  Read the documentation for facet_wrap(). What arguments can you use to control how many rows and columns appear in the output?

> __Answer__:

?facet_wrap: the arguments are nrow and ncol.

> __Question 4__: 

>  What does the scales argument to facet_wrap() do? When might you use it?

> __Answer__:

The default scale is fixed. This is flexible through "free" and "free_x" / "free_y" to explore across individual pattersns, then "free" scale would be more appropriate.

### Exercises 2.6

> __Question 1__: 

> What's the problem with the plot created by ggplot(mpg, aes(cty, hwy)) + geom_point()? Which of the geoms described above is most effective at remedying the problem?

> __Answer__:

Shows a simply comparison against to variables. Improvement could be to facet by class and add a smooth line to visualize the relationship.

```{r ch_2_6.1_1}

ggplot(mpg, aes(cty, hwy)) +
	geom_point() +
    geom_smooth() +
    facet_wrap(~class)

```

There is also an issue of overplotting. These could help:

```{r ch_2_6.1_2}

ggplot(mpg, aes(cty, hwy)) +
	geom_jitter() +
    geom_smooth()

```

> __Question 2__: 

> One challenge with ggplot(mpg, aes(class, hwy)) + geom_boxplot() is that the ordering of class is alphabetical, which is not terribly useful. 
How could you change the factor levels to be more informative? Rather than reordering the factor by hand, you can do it automatically based on the data: 
ggplot(mpg, aes(reorder(class, hwy), hwy)) + geom_boxplot(). What does reorder() do? Read the documentation.

> __Answer__:


```{r ch_2_6.2}

ggplot(mpg, aes(class, hwy)) + geom_boxplot()

ggplot(mpg, aes(reorder(class, hwy, fun = mean), hwy)) + geom_boxplot()

```

> __Question 3__: 

> Explore the distribution of the carat variable in the diamonds dataset. 
What binwidth reveals the most interesting patterns?

> __Answer__:


```{r ch_2_6.3}

ggplot(diamonds, aes(carat)) + geom_histogram()

ggplot(diamonds, aes(carat, color = cut)) +
    geom_freqpoly()

bins = 200
ggplot(diamonds, aes(x = carat)) +
	geom_histogram(bins = bins) +
	scale_x_continuous(limits = c(0, 2), expand = c(0, 0), breaks = seq(0, 2, 0.1))

```

> __Question 4__: 

> Explore the distribution of the price variable in the diamonds data. 
How does the distribution vary by cut?

> __Answer__:


```{r ch_2_6.4}

bins = 200
ggplot(diamonds, aes(x = price)) +
    geom_histogram(bins = bins) +
    facet_wrap(~cut, scales = "free")

ggplot(diamonds, aes(price, color = carat)) +
    geom_histogram(bins = bins) +
    facet_wrap(~cut, ncol = 2, scales = "free")

ggplot(diamonds, aes(x = price, y = ..density.., color = cut)) +
	geom_freqpoly(binwidth = 200)

```


> __Question 5__: 

> You now know (at least) three ways to compare the distributions of subgroups: 
    geom_violin(), 
    geom_freqpoly() and the colour aesthetic, or 
    geom_histogram() and facetting. 

    What are the strengths and weaknesses of each approach? What other approaches could you try?

> __Answer__:


```{r ch_2_6.5}


```

> __Question 6__: 

> You now know (at least) three ways to compare the distributions of subgroups: 
    geom_violin(), 
    geom_freqpoly() and the colour aesthetic, or 
    geom_histogram() and facetting. 

> What are the strengths and weaknesses of each approach? What other approaches could you try?

> __Answer__:


```{r ch_2_6.6}


```

> __Question 6__: 

> Using the techniques already discussed in this chapter, come up with three ways to visualise a 2d categorical distribution. 
Try them out by visualising the distribution of model and manufacturer, trans and class, and cyl and trans.

> __Answer__:


```{r ch_2_6.7}



```